---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# textrecipes

[![Travis build status](https://travis-ci.org/EmilHvitfeldt/textrecipes.svg?branch=master)](https://travis-ci.org/EmilHvitfeldt/textrecipes)
[![Coverage status](https://codecov.io/gh/EmilHvitfeldt/textrecipes/branch/master/graph/badge.svg)](https://codecov.io/github/EmilHvitfeldt/textrecipes?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/textrecipes)](http://cran.r-project.org/web/packages/textrecipes)
[![Downloads](http://cranlogs.r-pkg.org/badges/textrecipes)](http://cran.rstudio.com/package=textrecipes)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

## Introduction

**textrecipes** contains extra steps for the [`recipes`](http://cran.rstudio.com/package=recipes) package for preprocessing text data. 

## Installation

textrecipes is not avaliable from [CRAN](https://CRAN.R-project.org) yet. But the development version can be downloaded with: 

```{r installation, eval=FALSE}
require("devtools")
install_github("tidymodels/textrecipes")
```

## Example

In the following example we will go through the steps needed to convert a character variable to the TF-IDF of its tokenized words after removing stopwords and limeting ourself to only the 500 most used words. We will be conduction this preprosession on the variable `essay0`.

```{r, message=FALSE}
library(recipes)
library(textrecipes)

data(okc_text)

okc_rec <- recipe(~ ., data = okc_text) %>%
  add_role(contains("essay"), new_role = "textual") %>%
  step_tokenize(has_role("textual")) %>% # Tokenizes to words by default
  step_stopwords(has_role("textual")) %>% # Uses the english snowball list by default
  step_tokenfilter(has_role("textual"), max_tokens = 100) %>%
  step_tfidf(has_role("textual"))
   
okc_obj <- okc_rec %>%
  prep(training = okc_text)
   
str(bake(okc_obj, okc_text), list.len = 15)
```


## Type chart

**textrecipes** includes a little departure in design from **recipes** in the sense that it allows some input and output to be in the form of list columns. To avoind confusion here is a table of steps with their expected input and output respectively. Notice how you need to end with numeric for future analysis to work.

| Step               | Input       | Output      | Status  |
|--------------------|-------------|-------------|---------|
| `step_tokenize`    | character   | list-column | Done    |
| `step_untokenize`  | list-column | character   | Done    |
| `step_stem`        | list-column | list-column | Done    |
| `step_stopwords`   | list-column | list-column | Done    |
| `step_tokenfilter` | list-column | list-column | Done    |
| `step_tfidf`       | list-column | numeric     | Done    |
| `step_tf`          | list-column | numeric     | Done    |
| `step_texthash`    | list-column | numeric     | Done    |
| `step_word2vec`    | character   | numeric     | TODO    |

(TODO = Yet to be implemented, bug = correnctly not working, working = the step works but still not finished i.e. missing document/tests/arguemnts, done = finished)

This means that valid sequences includes

```{r, eval=FALSE}
recipe(~ ., data = data) %>%
  step_tokenize(text) %>%
  step_stem(text) %>%
  step_stopwords(text) %>%
  step_topwords(text) %>%
  step_tf(text)

# or

recipe(~ ., data = data) %>%
  step_tokenize(text) %>%
  step_stem(text) %>%
  step_tfidf(text)
```

