---
title: "Working with n-grams"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with n-grams}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(textrecipes)
library(tokenizers)
```

If you want to use n-grams with textrecipes you have 2 options:

- Use a tokenizer in `step_tokenize()` that tokenizes to n-grams.
- Tokenize to words with `step_tokenize()` and use `step_ngram()` to turn them into n-grams.

Both of these methods come with pros and cons so it will be worthwhile for you to be aware of both.

before we get started let's make sure we are on the same page of what we mean when we are talking about n-grams. We normally tokenize our text into words, which we can do with `tokenize_words()` from the tokenizers package (this is the default engine and token for `step_tokenize()` in textrecipes)

```{r}
abc <- c("The Bank is a place where you put your money;",
         "The Bee is an insect that gathers honey.")

tokenize_words(abc)
```

N-grams are a contiguous sequence of n tokens. So to get 2-gram (or bigrams as they are also called) we can use the `tokenize_ngrams()` function to get them

```{r}
tokenize_ngrams(abc, n = 2)
```

Notice how the words appear in multiple n-grams as the window slides across them.
And why changing the `n` argument we can any kind of n-gram (notice how `n = 1` is the special case of tokenizing to words).

```{r}
tokenize_ngrams(abc, n = 3)

tokenize_ngrams(abc, n = 1)
```

It can also be beneficial to specify a delimiter between the tokens in your n-gram.

```{r}
tokenize_ngrams(abc, n = 3, ngram_delim = "_")
```

# Only using `step_tokenize()`

The first methods work by using n-gram `token` from one of the built-in engine in `step_tokenize()` to get a full list of available tokens type `?step_tokenize()` and go down to `Details`.
We can use the `token="ngrams"` along with `engine = "tokenizers"`(the default) to tokenize to n-grams.
We finish this `recipe()` with `step_tokenfilter()` and `step_tf()`.
The filtering doesn't do anything to the data of this size but it is a good practice to use `step_tokenfilter()` before using `step_tf()` or `step_tfidf()` to control the size of the resulting data.frame.

```{r}
abc_tibble <- tibble(text = abc)

rec <- recipe(~ text, data = abc_tibble) %>%
  step_tokenize(text, token = "ngrams") %>%
  step_tokenfilter(text) %>%
  step_tf(text)

abc_ngram <- rec %>%
  prep() %>%
  juice()

abc_ngram

names(abc_ngram)
```

If you need to pass arguments to the underlying tokenizer function you can pass a named list to the `options` argument in `step_tokenize()`

```{r}
abc_tibble <- tibble(text = abc)

rec <- recipe(~ text, data = abc_tibble) %>%
  step_tokenize(text, token = "ngrams", options = list(n = 2, 
                                                       ngram_delim = "_")) %>%
  step_tokenfilter(text) %>%
  step_tf(text)

abc_ngram <- rec %>%
  prep() %>%
  juice()

abc_ngram

names(abc_ngram)
```

Lastly you can also supply a custom tokenizer to `step_tokenize()` using the `custom_token` argument.

```{r}
abc_tibble <- tibble(text = abc)

bigram <- function(x) {
  tokenizers::tokenize_ngrams(x, lowercase = FALSE, n = 2, ngram_delim = ".")
}

rec <- recipe(~ text, data = abc_tibble) %>%
  step_tokenize(text, custom_token = bigram) %>%
  step_tokenfilter(text) %>%
  step_tf(text)

abc_ngram <- rec %>%
  prep() %>%
  juice()

abc_ngram

names(abc_ngram)
```

Pros:

- Only uses 1 step
- Simple to use

Cons:

- Minimal flexibility, (`tokenizers::tokenize_ngrams()` don't let you control how the words are tokenized.)
- You are not able to tune the number of tokens in your n-gram

# Using `step_tokenize()` and `step_ngram()`

As of version 0.2.0 you can use `step_ngram()` along with `step_tokenize()` to gain higher control over how your n-grams are being generated.

```{r}
abc_tibble <- tibble(text = abc)

rec <- recipe(~ text, data = abc_tibble) %>%
  step_tokenize(text) %>%
  step_ngram(text, num_tokens = 3) %>%
  step_tokenfilter(text) %>%
  step_tf(text)

abc_ngram <- rec %>%
  prep() %>%
  juice()

abc_ngram

names(abc_ngram)
```

Now you are able to perform additional steps between the tokenization and the n-gram creation such as stemming the tokens.

```{r}
abc_tibble <- tibble(text = abc)

rec <- recipe(~ text, data = abc_tibble) %>%
  step_tokenize(text) %>%
  step_stem(text) %>%
  step_ngram(text, num_tokens = 3) %>%
  step_tokenfilter(text) %>%
  step_tf(text)

abc_ngram <- rec %>%
  prep() %>%
  juice()

abc_ngram

names(abc_ngram)
```

This also works great for cases where you need higher flexibility or when you want to use a more powerful engine such as **spacyr** that doesn't come with an n-gram tokenizer.

Furthermore the `num_tokens` argument is tunable with the **dials** and **tune** package.

Pros:

- Full flexibility
- Number of tokens is tunable

Cons:

- 1 Additional step is needed
